So, you've downloaded your very own copy of \Dumux and its dependencies.
You've run dunecontrol, and your example ``test$\_$dumux" not only complies,
but it even shows a nice simulation in paraview. 
Maybe you've read through parts of the handbook, and even started looking 
though the doxygen documentation. 
Well done. What now? \\ \linebreak

\textit{``How on earth is this going to help me solve my multi-(phase, component, 
scale, physics) flow and transport problems in porous media systems?''} you begin to wonder. 
Don't panic! In order to best ease our prospective users and developers into the
wonderful \Dumux simulation environment, we've prepared a \Dumux-course. 
This course is offered once a year over a period of 3 days at the University of Stuttgart.
If you're looking for information on attending, subscribe to the \Dumux mailing list 
and stay tuned for updates. \\
\url{https://listserv.uni-stuttgart.de/mailman/listinfo/dumux} \\ \linebreak

\textit{``But the course won't take place for another 6 months!"} and, 
\textit{``I want to start developing a numerical model of my challenging and 
	interesting process now!"}, you think. 
Not a problem. The course materials are all shared online in their own 
\Dumux project repository. A series of beginner-level exercises are explained 
such that you can see how a model is developed in \Dumux. As a teaser, we've
 also included a suite of examples from hot topics we're working on. Models
  exploring ``Coupling free flow and porous-media flow", ``Flow in fractured
   porous media" and ``Fluid-solid phase change" are all introduced.  \\ \linebreak
   
\textit{``Sounds great, but where is this material? I can't find it within
what I've downloaded."}, you question. 
The dumux course material is found online here: \\
\url{https://git.iws.uni-stuttgart.de/dumux-repositories/dumux-course} \\
In order to download this repository, which acts as an additional module to 
the \Dumux base, you can download an installation script with the following command:
\begin{lstlisting}[style=Bash]
$ wget https://git.iws.uni-stuttgart.de/dumux-repositories/dumux-course/raw/master/scripts/install.sh
\end{lstlisting}
This script will install \Dumux, it's Dune dependencies, and the \Dumux-course 
repository. Within the directory \Dumux-course there are a series of exercises 
and slides describing the previously described examples. \\ \linebreak

%TODO: Is this true? Did we update the dumux course to compatibility with 3.0?
At the moment, the \Dumux course is dependent on a certain version of
\Dumux. To make sure that this version of \Dumux is collected to your installation, 
check to see if the branch \textbf{dumux-course-2018}
is checked out within your dumux folder.
In the future, the \Dumux course will be updated to depend on
the specific release (3.0, 3.1) that was closest to the date of the course.