\section{Further Practice}
\label{tutorial-furtherpractice}

If there is a need for further practice, we refer here to the test problems, that
are already implemented in \Dumux. Several examples for coupled and decoupled models
can be found in the \texttt{test}-directory. An overview over the available tests
cases can be found on the class documentation \url{http://www.dumux.org/documentation.php}.
There you also find a \emph{feature-list} for the individual tests.

Another possibility to gain more experience with \Dumux is the \texttt{dumux-lecture} module, where 
a number of \Dumux applications are provided and explained. It is structured according to
the use of these applications in different lecture classes at the Department of Hydromechanics and
Modelling of Hydrosystems. The majority of applications belongs accordingly to the 
course in Multiphase Modelling (mm), while there are also some basic examples from the
courses Environmental Fluid Mechanics (efm) and Modelling of Hydrosystems (mhs). 
These applications contain were primarily designed to enhance the understanding of conceptualizing the
governing physical processes and their implementation in a numerical simulator. 
Different aspects of modelling multi-phase multi-component flow and transport processes are approached.
In the focus are questions like the assignment of boundary conditions, the choice of the 
appropriate physics for a given problem (which phases, which components), discretization issues,
time stepping, etc. 
You can find, for example, a comparison of different two-phase flow problems considering in
the more simple approach two immiscible fluids while components in both phases with interphase
mass transfer are considered in the more complex case.
All the scenarios and their physical background are explained in additional .tex-files,
which are provided in sub-directories names \emph{description}. The following test cases are 
contained:
\begin{itemize}
\item \texttt{Buckley-Leverett Problem - Classical porous media flow show case}
\item \texttt{$\text{CO}_2$ plume - The influence of the Gravitational Number}
\item \texttt{Column Xylene - A VEGAS experiment}
\item \texttt{Convective Mixing - Related to CO$_2$ storage}
\item \texttt{Fuel Cell}
\item \texttt{Heatpipe - A show case for two-phase two-component flow with heat fluxes}
\item \texttt{Heavy Oil - SAGD (steam assisted gravity drainage)}
\item \texttt{Henry Problem - A show case related to salt water intrusion}
\item \texttt{McWorther Problem - Classical porous media flow show case}
\item \texttt{NAPL Infiltration}
\item \texttt{Remediation Scenarios - For NAPL contaminated unsaturated soils}
\item \texttt{Groundwater - Simple groundwater flow case for MHS lecture}
\item \texttt{Different single/two-phase single/two-component problems - Example from lecture EFM}
\end{itemize}

Dumux-lecture can be obtained as follows:
\begin{lstlisting}[style=Bash]
$ git clone https://git.iws.uni-stuttgart.de/dumux-repositories/dumux-lecture.git
\end{lstlisting}
