\section[Quick start guide]{Quick start guide: The first run of a test application}\label{quick-start-guide}

The previous section showed how to install and compile \Dumux. This chapter shall give a very brief introduction how to run a first test application and how to visualize the first output files. Only the rough steps will be described here. More detailed explanations can be found in the tutorials in the following chapter.

\begin{enumerate}
 \item Go to the directory \texttt{/test}. There, various test application folders can be found. Let us consider as example \texttt{boxmodels/test{\_}2p}:
 \item Enter the folder \texttt{implicit/2p}. If everything was compiled correctly, there should be two executables \texttt{test{\_}cc2p} and \texttt{test{\_}box2p}. Otherwise, type \texttt{make test{\_}box2p} in order to compile the application \texttt{test{\_}box2p}. To run the simulation, type\\ 
 \texttt{./test{\_}box2p -parameterFile ./test\_box2p.input}\\
 into the console. The parameter \texttt{-parameterFile} specifies that all important parameters (like first timestep size, end of simulation and location of the grid file) can be found in a text file in the same directory  with the name \texttt{test\_box2p.input}.\\
 The same procedure can also be done for the \texttt{test{\_}cc2p}.
 \item The simulation starts and produces some .vtu output files and also a .pvd file. The .pvd file can be used to examine time series and summarizes the .vtu files. It is possible to stop a running application by pressing $<$Ctrl$><$c$>$.
 \item You can display the results using the visualization tool ParaView (or alternatively VisIt). Just type \texttt{paraview} in the console and open the .pvd file. On the left hand side, you can choose the desired parameter to be displayed.
\end{enumerate}
